---
title: "FinalProject"
author: "Faigah, Laila, Mrayam, Raghad"
date: "27/11/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###Project Name: Analyzing the Impact of a Country’s Economic Growth on Unemployment as Opposed to Other Factors and Their Effect

# Packages

```{r load-packages, message = FALSE}
library(tidyverse)
#library(randomForest)
library(dplyr)
#library(tseries)
library(ggplot2)
#library(imputeTS)
#library(forecast)
```

# Data

```{r load-data, message = FALSE}
NetExports <- read_csv("NetExports.csv")
InflationRates<-read_csv("InflationRates.csv")
GDP<-read_csv("GDP.csv")
UnemploymentRates<-read_csv("UnemploymentRates.csv")
CurrencyExchangeRates<-read_csv("CurrencyExchangeRates.csv")
GDPX<-read_excel("GDP.xlsx")
```

# Merging Data

```{r pressure, echo=FALSE}
plot(pressure)
```

##Discriptive Analysis

```{r hello}
View(data)
str(data)
boxplot(GDPX$`Data Source`)

```

#Handling Missing Values

```{r}
data1<-na.omit(data)
str(data1)

```

##Automate the process of recommendations and come up with a model to predict 
#the recommendation of _____ (in binary output- 0 or 1) accordingly.
#We Can use Naive Bayes,SVM, Decision Tree and Random Forest and test the best accuracy

```{r}
#1. Naive Bayes----------71.49% Accuracy
naive_bayes<-naiveBayes(Recommendation~.,data = data1)
summary(naive_bayes)

Predictions<-predict(naive_bayes,data1)
Predictions
table(Predictions,data1$Recommendation)

#Accuracy of Model
table_mat<-table(Predictions,data1$Recommendation)
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
accuracy_Test

#2. Support Vector Machine-------63.51% Accuracy
svms<-svm(Recommendation~.,data = data1,method = 'class')
summary(svms)

Predictions<-predict(svms,data1,type='class')
Predictions
table(Predictions,data1$Recommendation)

#Accuracy of Model
table_mat<-table(Predictions,data1$Recommendation)
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
accuracy_Test

#3. Decision Tree---------79.44% Acuraccy
tree<-rpart(Recommendation~.,data = data1,method = 'class')
summary(tree)

Predictions<-predict(tree,data1,type = 'class')
Predictions
table(Predictions,data1$Recommendation)

#Accuracy of Model
table_mat<-table(Predictions,data1$Recommendation)
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
accuracy_Test

#4. Random Forest---------100% Accuracy
forest<-randomForest(x = data1, y = data1$Recommendation,ntree =800)# build model
summary(forest)

Predictions<-predict(forest,data1)
Predictions
table(Predictions,data1$Recommendation)

#Accuracy of Model
table_mat<-table(Predictions,data1$Recommendation)
accuracy_Test <- sum(diag(table_mat)) / sum(table_mat)
accuracy_Test


```

#Time Series Analysis using ARIMA

```{r}
data1<-na.omit(data)
str(data1)

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

hellohellohellohellohellohellohello