---
title: "Analyzing the Impact of a Country’s Economic Growth on Unemployment as Opposed to Other Factors and Their Effects"
author: "Faigah Bajammal, Laila Bakhashab, Mrayam Alnahdi, Raghad Zarei"
date: "14th Dec 2021"
output: html_document
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

```{r load-packages, message = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
install.packages(arsenal)
install.packages("caret")
library(caret)
install.packages("RANN")
library(RANN)  # required for knnImpute
library(ada)
install.packages("ada")
install.packages("readxl")
library(readxl)
library(ggridges)
install.packages("ggridges")
library(ggplot2)
library(fivethirtyeight)
library(broom) 
install.packages("readr")
library(readr)
```

# Data

```{r load-data, message = FALSE}

NetExports <- read_csv("NetExports.csv")
InflationRates<-read_csv("InflationRates.csv")
CurrencyExchangeRates<-read_csv("CurrencyExchangeRates.csv")
unemploymentRate<-read_csv("unemploymentRate.csv")
ExchangeRate<-read_csv("ExchangeRate.csv")
GDP<-read_excel("GDP.xls")

```

##Discriptive Analysis

#Ex-1 Which country's exchange rate was impacted the most due to the us great recession through 2006-2009, which country was least impacted? How did unemployment look in both countries? Was there a relationship? 

```{r Ex-1}

exchange_greatRec_high <- ExchangeRate %>% filter(TIME==2009) %>% group_by(LOCATION)%>% summarize(max=max(Value),TIME=mean(TIME)) %>% arrange(desc(max))%>%slice(1, n())

exchange_pre_greatRec <- ExchangeRate[c(1,6,7)] %>% 
filter(LOCATION=='IDN'| LOCATION=='GBR', TIME == 2006)

#exDiff <- merge(exchange_greatRec_high,exchange_pre_greatRec, by= c("LOCATION","TIME","Value"))
#merge(x = exchange_greatRec_high, y = exchange_pre_greatRec)
#merge(x = exchange_greatRec_high, y = exchange_pre_greatRec, by = c("LOCATION","TIME","Value"))
#exDiff <- exchange_greatRec_high %>%
#inner_join(exchange_pre_greatRec, by = c("LOCATION","TIME"))

exDiff <- merge(exchange_greatRec_high ,exchange_pre_greatRec, by ="Value", all = TRUE)
exDiff   


unemploymentRate %>% filter(Code %in% c("IDN" , "GBR") ) %>%  filter(Year %in% c(2006,2007,2008,2009) ) %>%
ggplot(unemploymentRate, mapping = aes(x = Unemployment ,col="blue", y = Code, label = Year)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")

```

#Ex-2 OCED members in unemployment (Was the NetExports affected by the unemplyment rates in OCED countries? 1. we have to filter the OCED countires in unemplyment 2. Using a Graph check how netexports was affected.

```{r ex-2}

OECD_exp <- NetExports[c(1,3,4,6,7)] %>% filter(TIME>=1995, Country == "OECD", SUBJECT=="EXP") %>% rename(net_Trade = Value) %>% rename(Year = TIME)

# filter Country=='OECD'
oecd_unemployment <-unemploymentRate[c(1,3,4)] %>% filter( Year>=1995) %>% filter(Country == 'OECD')

OECD_compare <- inner_join(oecd_unemployment,OECD_netTrade, by = "Year")

ggplot(OECD_compare,
       mapping =aes(x=net_Trade, y=Year, size = Unemployment))+
  geom_point(alpha = 0.90)+
  labs(x = "Exports", y = "Year",
       size = "Unemployment",
       title = "Relationship between Unemployment and exports of OECD countries")

```

#Ex-3 In 1980, 1998, and 2019, the world showed the most inflation rates, is it correlated with GDP?

```{r ex-3}

plot(GDPX$'1980', InflationRates$'1980', type = "l")
plot(GDPX$'1998', InflationRates$'1998', type = "b")
plot(GDPX$'2019', InflationRates$'2019', type = "o")

plot(GDPX$'1998', InflationRates$'1998',
     main = "Correlation of GDP and Inflation Rates",
     xlab = "GDP",
     ylab = "Inflation Rates",
     type = "b",
     col="blue") 

lines(GDPX$'1980', InflationRates$'1980',
      lty = "dotted",
      col="green")

lines(GDPX$'2019', InflationRates$'2019',
      lty = "dashed",
      col="red")

labels <- c("1980", "1998", "2019")

legend("topright", legend=labels,
       cex = 0.6,
       inset = 0.01,
       pch = NULL,
       lty = c(1,2,3))

```

##Automate the process of recommendations and come up with a model to predict 
#the recommendation of _____ (in binary output- 0 or 1) accordingly.
#We Can use Naive Bayes,SVM, Decision Tree and Random Forest and test the best accuracy

```{r}

#Handling Missing Values
CER<-na.omit(CurrencyExchangeRates)
str(CER)
```

#Build Regression Model

```{r}
#regression <-lm(GDP$'2020'~unemploymentRate$'2020')
#regression
#summary(regression)
#Predictions<-predict(regression,unemploymentRate)
#Predictions
```
# Data Preparation and Preprocessing

```{r, echo=FALSE}
# Step 1: Get row numbers for the training data
trainRowNumbers <- createDataPartition(CER$'Saudi Arabian Riyal', p=0.8, list=FALSE)

# Step 2: Create the training  dataset
trainData <- CER[trainRowNumbers,]

# Step 3: Create the test dataset
testData <- CER[-trainRowNumbers,]

# Store X and Y for later use.
x = trainData[, 2:18]
y = trainData$'Saudi Arabian Riyal'

# Create the knn imputation model on the training data
CERKNN <- preProcess(trainData, method='knnImpute')
CERKNN #knn imputation acheived 


# Use the imputation model to predict the values of missing data points
trainData <- predict(CERKNN, newdata = trainData)
anyNA(trainData)

#One-Hot Encoding 
dummies_model <- dummyVars( `Saudi Arabian Riyal`~ ., data=trainData)

trainData_mat <- predict(dummies_model, newdata = trainData)

trainData <- data.frame(trainData_mat)

str(trainData)

#Feature Selection Using Recursive Feature Elimination 

#We built a ML model on a training dataset and estimate the feature importance on the test dataset while keeping priority to the most important variables, iterate through by building models of given subset sizes, that is, subgroups of most important predictors determined from step 1. Ranking of the predictors is recalculated in each iteration and show the model performances are compared across different subset sizes to arrive at the optimal number and list of final predictors.

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   verbose = FALSE)

#To know what models caret supports, we ran the following:
modelnames <- paste(names(getModelInfo()), collapse=',  ')
modelnames
```


#Selecting a Machine Learning Model

```{r load-data, message = FALSE}

# Train the model using randomForest and predict on the training data itself.
model_CER = train(Date2018.4.4 ~ ., data=trainData, method='null')
fitted <- predict(model_CER)
model_CER

#Prepare the test dataset and predict
# Step 1: Impute missing values 
testData2 <- predict(CERKNN, testData)  

# Step 2: Create one-hot encodings (dummy variables)
testData3 <- predict(dummies_model, testData2)

# Step 3: View
head(testData3[, 1:10])

# Train the model using adaboost
model_adaboost = train(Date2018.4.4 ~ ., data=trainData, method='adaboost', tuneLength=2, trControl = fitControl)
ymodel_adaboost

# Train the model using rf
model_rf = train(Date2018.4.4 ~ ., data=trainData, method='rf', tuneLength=5, trControl = fitControl)
model_rf

# Train the model using CER
model_xgbDART = train(Date2018.4.4 ~ ., data=trainData, method='xgbDART', tuneLength=5, trControl = fitControl, verbose=F)
model_xgbDART

# Train the model using CER
model_svmRadial = train(Date2018.4.4 ~ ., data=trainData, method='svmRadial', tuneLength=15, trControl = fitControl)
model_svmRadial

# Compare model performances using resample()
models_compare <- resamples(list(ADABOOST=model_adaboost, RF=model_rf, XGBDART=model_xgbDART, MARS=model_mars3, SVM=model_svmRadial))

# Summary of the models performances
summary(models_compare)

```

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

